---
title: "Week 7"
output: html_document
date: '2022-08-03'
---

Loading Packages

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(gridExtra)
library(readr)
library(readxl)
library(tidyverse)
library(here)
library(skimr)
library(janitor)
library(ggplot2)
```


Loading in Data

```{r message=FALSE, warning=FALSE}
accord_key <- read_csv(paste("accord_key.csv", sep = ""))
activitystatus <- read_csv(paste("activitystatus.csv", sep = ""),
                           col_types = "dcdcdcddd")

bloodpressure <- read_csv(paste("bloodpressure.csv", sep = ""))
hba1c <- read_csv(paste("hba1c.csv", sep = ""))
otherlabs <- read_csv(paste("otherlabs.csv", sep = ""))

cvdoutcomes <- read_csv("cvdoutcomes.csv")
hypoglycemia_events <- read_csv(paste("hypoglycemiaevents.csv", sep = ""))
hypoglycemia_time1st <- read_csv(paste("hypoglycemiatime1st.csv", sep = ""))
sae <- read_csv(paste("sae.csv", sep = ""))


f07_baseline <- read.csv("f07_baselinehistoryphysicalexam.csv")
microvascular_outcomes <- read_csv("microvascularoutcomes.csv")
lipids <- read_csv("lipids.csv")
```


Fixing AG dataset

```{r message=FALSE, warning=FALSE}
AG <- read_xlsx(paste("UCLA Samples-2022-07-19 Summary.xlsx"), 
                sheet = "Summary", skip = 7) %>% 
  rename(BSI_ID = `Sample ID`)
linkfile <- read_csv("zhou_7816_linkfile.csv")

AG <- AG %>% 
  full_join(linkfile, by = "BSI_ID") %>% 
  filter(!is.na(MASKID) ) %>% 
  rename(MaskID = MASKID,
         Box = `R2020 000036`)

BatchIDs <- c("R2020 000036", "R2020 000037", "R2020 000038", "R2020 000039", 
              "R2020 000040", "R2020 000041", "R2020 000042", "R2020 000043")
index = c(1, which(AG$Box %in% BatchIDs), 4001)
BatchID_col = rep(BatchIDs, diff(index))
AG <- AG %>% mutate(BatchID = BatchID_col) %>%
  relocate(BatchID, .before = Box) %>%
  relocate(MaskID, .after = BSI_ID)

AG_summary <- AG %>% group_by(Box)
```


## Correlation of 1,5 AG measured at 12 months and othe biomarkers at 12 months

### Fixing data sets to merge

```{r}
BMI <- f07_baseline %>%
  mutate(m = ht_cm/100) %>%
  select(m, MaskID, wt_kg) %>%
  mutate(BMI = wt_kg/m^2)

  
bloodpressure_12month <- bloodpressure %>%
  filter(Visit == "F12")
hba1c_12month <- hba1c %>%
  filter(Visit == "F12")
lipids_12_month <- lipids %>%
  filter(Visit == "F12")
gfr_12_months <- otherlabs %>% 
  filter(Visit == "F12") %>%
  select(MaskID, gfr)

AG_12_month <- AG %>%
  left_join(hba1c_12month, by = "MaskID") %>%
  left_join(bloodpressure_12month, by = "MaskID") %>% 
  left_join(lipids_12_month, by = "MaskID") %>%
  left_join(gfr_12_months, by = "MaskID") %>%
  left_join(BMI, by = "MaskID") %>%
  within(rm(hr, Visit.y, vldl, Visit.x)) 
```



## Joining baseline covarties to 1,5 AG measurements

```{r}
accord_key <- accord_key %>%
  mutate(glycemic_arm = ifelse(arm==1|arm==2|arm==5|arm==6,0,1),
         bp_arm = ifelse(arm==3 | arm ==1,1,0),
         lipid_arm = ifelse(arm==7 | arm == 5,1,0))


baseline_ualb <- otherlabs %>% filter(Visit == "BLR") %>% 
  mutate(kidney = ifelse(gfr < 45 | ualb > 300, 1, 0)) %>% 
  select(MaskID, -Visit, kidney)


tindpt <- AG %>% 
  left_join(accord_key, by = "MaskID" ) %>%
  left_join(f07_baseline, by = "MaskID") %>%
  select(MaskID,
         baseline_age,
         glycemic_arm, bp_arm, lipid_arm, female, raceclass, cvd_hx_baseline,
         yrsdiab, hartfail, eyedisea, smokelif) %>%
  left_join(baseline_ualb, by = "MaskID")
```


```{r}
Full_AG <- AG_12_month %>%
  left_join(tindpt, by = "MaskID")
```


### Two level variable

```{r}
twolevel_AG <- Full_AG %>% 
  mutate(BatchID = recode(BatchID, "R2020 000036" = "outliers", 
                          "R2020 000037" = "outliers", "R2020 000038" = "38-43",
                          "R2020 000039" = "38-43",
                          "R2020 000040" = "38-43",
                          "R2020 000041" = "38-43",
                          "R2020 000042" = "38-43",
                          "R2020 000043" = "38-43",))

```

### Plotting biomarkers at 12 months with 1,5 AG

**BMI**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BMI, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```

**gfr**
```{r}
twolevel_AG %>%
  ggplot(aes(x = gfr, y = Mean)) +
  geom_jitter()+
  facet_wrap(~ BatchID)
```

**sbp**
```{r}
twolevel_AG %>%
  ggplot(aes(x = sbp, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```

**dbp**
```{r}
twolevel_AG %>%
  ggplot(aes(x = dbp, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```

**chol**
```{r}
twolevel_AG %>%
  ggplot(aes(x = chol, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```

**trig**
```{r}
twolevel_AG %>%
  ggplot(aes(x = trig, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```

**ldl**
```{r}
twolevel_AG %>%
  ggplot(aes(x = ldl, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```

**hdl**
```{r}
twolevel_AG %>%
  ggplot(aes(x = hdl, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```


### Baseline covarties plotting with 1,5 AG

**glycemic arm**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ glycemic_arm)
```

**bp arm**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ bp_arm)
```

**lipid arm**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ lipid_arm)
```

**female/male**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ female)
```

**raceclass**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ raceclass )
```

**cvd hx baseline**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ cvd_hx_baseline)
```

**eyedisea**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ eyedisea)
```

**smokelif**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ smokelif )
```

**kidney**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ kidney)
```

**hartfail**
```{r}
twolevel_AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter() +
  facet_wrap( ~ hartfail)
```

**yrsdiab**
```{r}
twolevel_AG %>%
  ggplot(aes(x = yrsdiab, y = Mean)) +
  geom_jitter() +
  facet_wrap(~BatchID)
```

**baseline age**
```{r}
twolevel_AG %>%
  ggplot(aes(x = baseline_age, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```

Print(Width = Inf)

