---
title: "Week 4 homework"
output: html_document
date: '2022-07-15'
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(skimr)
library(janitor)
library(ggplot2)
```

**Read in data**
```{r message=FALSE}
bloodpressure <- read_csv("bloodpressure.csv")
accord_key <- read_csv("accord_key.csv")
bp_hba1c <- read_csv("hba1c.csv")
```

### Brief description of datasets 

All three data sets contain MaskID which represents the number given to a patient. Each three of these data sets contain 10,251 obs which is the number of individuals.

Number of females #3952
Number of males  #6299
```{r echo=TRUE}
accord_key %>%
  group_by(female) %>%
  summarize(number = n(),
percent = 100 *number / 10251)
```

- 1953 Black, 19.1%
- 6393 White, 62.4%
- 737 Hispanic, 7.2%
- 1168 Other, 11.4%

### Mean, median, sd for BP/blood sugar measures per individual
**sbp**
```{r echo=TRUE, message=TRUE}
new_sbp <- bloodpressure %>%
  group_by(MaskID) %>%
  summarise(max = max(sbp),
            mean = mean(sbp),
            median = median(sbp),
            sd = sd(sbp),
            IQR = IQR(sbp))
```

**hba1c**
```{r}
new_hba1c <- bp_hba1c %>%
  group_by(MaskID) %>%
  summarise(max = max(hba1c),
            mean = mean(hba1c),
            median = median(hba1c),
            sd = sd(hba1c),
            IQR = IQR(hba1c))
```

### Distributions for baseline SBP, DBP, and HbA1c

```{r echo=TRUE}
#Baseline sbp and dbp
bp_basline <- bloodpressure%>%
  filter(Visit == "BLR")
#Baseline hba1c 
hba1c_baseline <- bp_hba1c %>%
  filter(Visit == "BLR")

#Baseline sbp, dbp, and hba1c
baseline_merge <- inner_join(bp_basline, hba1c_baseline)

#Baselines merged with demo info
b_merge2 <- inner_join(baseline_merge, accord_key, by = "MaskID")
```



### Mean, Median, Mode, SD of baseline SBP, DBP, and HbA1c for the whole population

```{r echo=TRUE, message=TRUE, paged.print=TRUE}
SBP_mean_sd_iqr_baseline <- b_merge2 %>%
summarise(max = max(sbp),
            mean = mean(sbp),
            median = median(sbp),
            sd = sd(sbp))

DBP_mean_sd_iqr_baseline_SBP <- b_merge2 %>%
  summarise(max = max(dbp),
            mean = mean(dbp),
            median = median(dbp),
            sd = sd(dbp))

HBA1C_mean_sd_iqr_baseline_SBP <- b_merge2 %>%
  summarise(max = max(hba1c),
            mean = mean(hba1c),
            median = median(hba1c),
            sd = sd(hba1c))
```

- sbp
results in a mean of 136.3, median of 135, and sd of 17.11545

- dbp
results in a mean of 74.9, median of 75, and sd of 10.66146

- hba1c
results in a mean of 8.3, median of 8.1, and sd of 1.057004

###  mean (SD) and median (IQR) of baseline SBP, DBP, and HbA1c for male/female 
```{r}
#sbp
b_merge2 %>%
  group_by(female) %>%
  summarise(max = max(sbp),
            mean = mean(sbp),
            median = median(sbp),
            sd = sd(sbp),
            IQR = IQR(sbp))
#dbp
b_merge2 %>%
  group_by(female) %>%
  summarise(mean = mean(dbp),
            median = median(dbp),
            sd = sd(dbp),
            IQR = IQR(dbp))

#hba1c
b_merge2 %>%
  group_by(female) %>%
  summarise(mean = mean(hba1c),
            median = median(hba1c),
            sd = sd(hba1c),
            IQR = IQR(hba1c))
```

### Mean (SD) and median (IQR) of baseline SBP, DBP, and HbA1c for different ethnic groups

```{r}
#hba1c
b_merge2 %>%
  group_by(raceclass) %>%
  summarise(mean = mean(hba1c),
            median = median(hba1c),
            sd = sd(hba1c),
            IQR = IQR(hba1c))
#dbp
b_merge2 %>%
  group_by(raceclass) %>%
  summarise(mean = mean(dbp),
            median = median(dbp),
            sd = sd(dbp),
            IQR = IQR(dbp))
#sbp
b_merge2 %>%
  group_by(raceclass) %>%
  summarise(mean = mean(sbp),
            median = median(sbp),
            sd = sd(sbp),
            IQR = IQR(sbp))
```


### Calculating the mean for SBP over all individuals 


```{r}
mean_sbp_bloodpress <- bloodpressure %>% 
  group_by(Visit) %>%
  summarize(meansbp = mean(sbp))

```

### Plotting the mean profile plot for sbp (mfp)
```{r}
mean_sbp_bloodpress %>% 
  ggplot(aes(x = Visit, y = meansbp)) +
  geom_jitter() +
  coord_flip()


```

### Plotting the mean profile for hba1c

```{r}
mean_hba1c_bp_hba1c <- bp_hba1c %>%
  group_by(Visit) %>%
  summarise(mean = mean(hba1c))
```

```{r}
mean_hba1c_bp_hba1c %>% 
  ggplot(aes(x = Visit, y = mean)) +
  geom_jitter() +
  coord_flip()
```

### Plotting mean profile for dbp

```{r}
mean_dbp_bloodpressure <- bloodpressure %>% 
  group_by(Visit) %>%
  summarise(mean_dbp = mean(dbp))

```

```{r warning=FALSE}
mean_dbp_bloodpressure %>%
  ggplot(aes(x = Visit, y = mean_dbp), na.rm == TRUE) +
  geom_jitter() +
  coord_flip()

```


### Loading in data

```{r message=FALSE, warning=FALSE}
accord_key <- read_csv(paste("accord_key.csv", sep = ""))
activitystatus <- read_csv(paste("activitystatus.csv", sep = ""),
                           col_types = "dcdcdcddd")

bloodpressure <- read_csv(paste("bloodpressure.csv", sep = ""))
hba1c <- read_csv(paste("hba1c.csv", sep = ""))
otherlabs <- read_csv(paste("otherlabs.csv", sep = ""))

cvdoutcomes <- read_csv("cvdoutcomes.csv")
hypoglycemia_events <- read_csv(paste("hypoglycemiaevents.csv", sep = ""))
hypoglycemia_time1st <- read_csv(paste("hypoglycemiatime1st.csv", sep = ""))
sae <- read_csv(paste("sae.csv", sep = ""))

```


```{r}
accord_key <- accord_key %>% 
  mutate(bp_arm = ifelse(arm == 1 | arm == 3, 
                         "intensive BP", "standard BP or not in BP arm"),
         glycemic_arm = ifelse(arm == 3 | arm == 4 | arm == 7 | arm == 8, 
                               "intensive glycemia", "standard glycemia"),
         glycemic_arm = factor(glycemic_arm, levels = 
                                 c("standard glycemia", "intensive glycemia")))

```

```{r}
trend1 <- accord_key %>%
  left_join(bloodpressure, by = "MaskID") %>%
  left_join(hba1c, by = c("MaskID", "Visit")) %>% 
  left_join(activitystatus, by = c("MaskID", "Visit")) %>%
  select(MaskID, Visit, days_from_baseline, sbp, dbp, hba1c, arm, bp_arm, glycemic_arm) %>% 
  filter(Visit != "EXIT")


table(trend1$glycemic_arm[trend1$Visit=="BLR"])
```

```{r}
trend_med <- trend %>% 
  group_by(Visit, glycemic_arm) %>% 
  summarize(median = median(hba1c, na.rm = TRUE),
            Q1 = ifelse(Visit == "BLR" | 
                          Visit == "F12" | Visit == "F24" | 
                          Visit == "F36" | Visit == "F48" | Visit == "F60" | Visit == "F72",
                        quantile(hba1c, 0.25, na.rm = TRUE), 0),
            Q3 = ifelse(Visit == "BLR" | 
                          Visit == "F12" | Visit == "F24" | 
                          Visit == "F36" | Visit == "F48" | Visit == "F60" | Visit == "F72",
                        quantile(hba1c, 0.75, na.rm = TRUE), 0)) %>% 
  filter(!is.na(median)) %>%
  slice_head()



trend_med

table(trend$glycemic_arm[trend$Visit %in% c("BLR", "F12", "F24", "F36", "F48", "F60", "F72", "F84")],
      trend$Visit[trend$Visit %in% c("BLR", "F12", "F24", "F36", "F48", "F60", "F72", "F84")])

table(trend$glycemic_arm[trend$days_from_baseline <= 48], 
      trend$days_from_baseline[trend$days_from_baseline <= 48])

table(accord_key$glycemic_arm)

trend %>% 
  filter(Visit == "BLR", glycemic_arm == "standard glycemia") 






### Reproduce the output


trend_med %>% 
  group_by(Visit, glycemic_arm) %>% 
  ggplot(aes(x = Visit, y = median, shape = glycemic_arm, color = glycemic_arm)) +
  geom_point(position = position_dodge(width = -0.5)) +
  geom_line(aes(group = glycemic_arm)) +
  geom_errorbar(aes(ymin = Q1, ymax = Q3), width = .4, position = position_dodge(width = -0.5)) +
  theme(legend.position = "bottom") + 
  labs(title = "Median Glycated Hemoglobin Levels at Each Study Visit",
       x = "Years", y = "Glycated Hemoglobin (%)") +
  scale_y_continuous(limits = c(6, 9))





```

