---
title: "Week 2/3"
output: html_document
date: '2022-08-22'
---

This week introduced me to tidying and analyzing data. I continued some of the chapters from week 1. One of the important things I learned was the pipe "%>%"

```{r}
# load packages ----

library(tidyverse)
library(here)
library(skimr)
library(janitor)
library(ggplot2)

```

```{r}
beaches <- read_csv(here("data1", "sydneybeaches.csv"))


# exploring the data -----

View(beaches)
dim(beaches)
str(beaches)
glimpse(beaches)
head(beaches)
tail(beaches)
summary(beaches)
skim(beaches)

```

```{r}
# tidying columns-----------

glimpse(beaches)

select_all(beaches, toupper)
select_all(beaches, tolower)

cleanbeaches <- clean_names(beaches)

names(cleanbeaches)

# for rename use newname = oldname
cleanbeaches <- rename(cleanbeaches, beachbugs = enterococci_cfu_100ml)


```

```{r}
# select a subset of columns
select(cleanbeaches, council, site, beachbugs, everything())

# pipe %>%

cleanbeaches <- beaches %>%
  clean_names() %>%
  rename(beachbugs = enterococci_cfu_100ml) 

write_csv(cleanbeaches, "cleanbeaches.csv")

```

Summarry Statistics
```{r}
# sorting and filtering ----

# which beach has the most extreme levels of bugs

worstbugs <- cleanbeaches %>% arrange(desc(beachbugs))

worstcoogee <- cleanbeaches %>% filter(site == "Coogee Beach") %>%
  arrange(-beachbugs)

# lets compare max bug value across different beaches

cleanbeaches %>%
  filter(site %in% c("Coogee Beach", "Bondi Beach")) %>%
  arrange(-beachbugs)

#group_by and summarise -------
cleanbeaches %>%
  filter(site %in% c("Coogee Beach", "Bondi Beach")) %>%
  group_by(site) %>%
  summarise(maxbug = max(beachbugs, na.rm = TRUE),
           meanbugs = mean(beachbugs, na.rm = TRUE),
           medianbugs = median(beachbugs, na.rm = TRUE),
           sdbugs = sd(beachbugs, na.rm = TRUE))

```

```{r}
# lets compare councils

cleanbeaches %>% distinct(council)

councilbysite <- cleanbeaches %>% 
  group_by(council, site) %>%
  summarise(meanbugs = mean(beachbugs, na.rm = TRUE),
            medianbugs = median(beachbugs, na.rm = TRUE))


```

```{r}
# compute new variable -----

glimpse(cleanbeaches)

testdate <- cleanbeaches %>% separate(date, c("day", "month", "year"), remove = FALSE)

cleanbeaches %>% unite(council_site, council:site, remove = FALSE )

# use mutate to transform the beachbugs data

summary(cleanbeaches)

cleanbeaches %>% mutate(logbeachbugs = log(beachbugs))

```

```{r}
# use mutate to computer new numeric variable

cleanbeaches %>% mutate(beachbugsdiff = beachbugs - lag(beachbugs))

# use mutate to compute new logical data

cleanbeaches %>% mutate(buggier = beachbugs > mean(beachbugs, na.rm = TRUE))

meanbugs = mean(cleanbeaches$beachbugs, na.rm = TRUE)

cleanbeaches_new <- cleanbeaches %>% 
  separate(date, c("day", "month", "year"), remove = FALSE) %>%
  mutate(logbeachbugs = log(beachbugs))%>%
  mutate(beachbugsdiff = beachbugs - lag(beachbugs)) %>%
  mutate(buggier_all = beachbugs > mean(beachbugs, na.rm = TRUE)) %>%
  group_by(site) %>%
  mutate(buggier_site = beachbugs > mean(beachbugs, na.rm = TRUE))

```

```{r}
# write clean data to .csv -----

write_csv(cleanbeaches_new, here("data1", "cleanbeaches_new.csv"))

```

## Modern Drive

```{r}
# load packages -------

library(ggplot2)
library(dplyr)
library(nycflights13)
library(knitr)
library(moderndive)
library(readr)
library(tidyr)
library(fivethirtyeight)

# read in Data ----

View(airlines)
View(flights)
View(alaska_flights)
glimpse(flights)
kable(airlines)
airlines$name
glimpse(airports)
?flights

```

```{r}
# Scatter Plot -----

alaska_flights <- flights %>% 
  filter(carrier == "AS")

ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) + 
  geom_point()

ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) + 
  geom_point(alpha = 0.2)

ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) + 
  geom_jitter(width = 30, height = 30)


```


```{r}
# Line Graphs -----
# usually uses time as the dependent variable
View(weather)
View(early_january_weather)
glimpse(weather)
?weather

early_january_weather <- weather %>% 
  filter(origin == "EWR" & month == 1 & day <= 15)

ggplot(data = early_january_weather, mapping = aes(x = time_hour, y = temp)) +
  geom_line()

```

```{r}
# Histograms/Bargraphs -----

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram()

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(color = "white")

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(color = "white", fill = "steelblue")

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(bins = 40, color = "white")

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(binwidth = 10, color = "white")


```

```{r}
# Facets -----

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(binwidth = 5, color = "white") +
  facet_wrap(~ month)

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(binwidth = 5, color = "white") +
  facet_wrap(~ month, nrow = 4)


```

```{r}
# Box plots -----

ggplot(data = weather, mapping = aes(x = month, y = temp)) +
  geom_boxplot()

ggplot(data = weather, mapping = aes(x = factor(month), y = temp)) +
  geom_boxplot()

```

```{r}
# Barplots -----

ggplot(data = flights, mapping = aes(x = carrier)) +
  geom_bar()

ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar()

ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar(position = "dodge")

ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar(position = position_dodge(preserve = "single"))

ggplot(data = flights, mapping = aes(x = carrier)) +
  geom_bar() +
  facet_wrap(~ origin, ncol = 1)
```




# SUMMARY ----
"1 Scatterplot	Relationship between 2 numerical variables	geom_point()	
2	Linegraph	Relationship between 2 numerical variables	geom_line()	Used when 
there is a sequential order to x-variable, e.g., time
3	Histogram	Distribution of 1 numerical variable	geom_histogram()	Facetted 
histograms show the distribution of 1 numerical variable split by the values of 
another variable
4	Boxplot	Distribution of 1 numerical variable split by the values of another 
variable	geom_boxplot()	
5	Barplot	Distribution of 1 categorical variable	geom_bar() when counts are not 
pre-counted, geom_col() when counts are pre-counted	Stacked, side-by-side, and 
faceted barplots show the joint distribution of 2 categorical variables"

```{r}
# Segment 1:
ggplot(data = flights, mapping = aes(x = carrier)) +
  geom_bar()

```

```{r}

# Segment 2:
ggplot(flights, aes(x = carrier)) +
  geom_bar()
alaska_flights <- flights %>% 
  filter(carrier == "AS")

portland_flights <- flights %>% 
  filter(dest == "PDX")
View(portland_flights)


```




### CHPT 3 

```{r}
btv_sea_flights_fall <- flights %>% 
  filter(origin == "JFK" & (dest == "BTV" | dest == "SEA") & month >= 10)
View(btv_sea_flights_fall)
     
not_BTV_SEA <- flights %>% 
  filter(!(dest == "BTV" | dest == "SEA"))
View(not_BTV_SEA)

many_airports <- flights %>% 
  filter(dest %in% c("SEA", "SFO", "PDX", "BTV", "BDL"))
View(many_airports)
```

```{r}
# Gives out NA values, the code below it excludes these NA values
summary_temp <- weather %>% 
  summarize(mean = mean(temp), std_dev = sd(temp))
summary_temp

summary_temp <- weather %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_temp


summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_monthly_temp


by_origin_monthly <- flights %>% 
  group_by(origin, month) %>% 
  summarize(count = n())
by_origin_monthly


weather <- weather %>% 
  mutate(temp_in_C = (temp - 32) / 1.8)

summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(mean_temp_in_F = mean(temp, na.rm = TRUE), 
            mean_temp_in_C = mean(temp_in_C, na.rm = TRUE))
summary_monthly_temp


flights <- flights %>% 
  mutate(gain = dep_delay - arr_delay)

gain_summary <- flights %>% 
  summarize(
    min = min(gain, na.rm = TRUE),
    q1 = quantile(gain, 0.25, na.rm = TRUE),
    median = quantile(gain, 0.5, na.rm = TRUE),
    q3 = quantile(gain, 0.75, na.rm = TRUE),
    max = max(gain, na.rm = TRUE),
    mean = mean(gain, na.rm = TRUE),
    sd = sd(gain, na.rm = TRUE),
    missing = sum(is.na(gain))
  )
gain_summary

ggplot(data = flights, mapping = aes(x = gain)) +
  geom_histogram(color = "white", bins = 20)

flights <- flights %>% 
  mutate(
    gain = dep_delay - arr_delay,
    hours = air_time / 60,
    gain_per_hour = gain / hours
  )

freq_dest <- flights %>% 
  group_by(dest) %>% 
  summarize(num_flights = n())
freq_dest

freq_dest %>% 
  arrange(desc(num_flights))

flights_joined <- flights %>% 
  inner_join(airlines, by = "carrier")
View(flights)
View(flights_joined)


flights_with_airport_names <- flights %>% 
  inner_join(airports, by = c("dest" = "faa"))
View(flights_with_airport_names)

named_dests <- flights %>%
  group_by(dest) %>%
  summarize(num_flights = n()) %>%
  arrange(desc(num_flights)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  rename(airport_name = name)
named_dests


flights_weather_joined <- flights %>%
  inner_join(weather, by = c("year", "month", "day", "hour", "origin"))
View(flights_weather_joined)

joined_flights <- flights %>% 
  inner_join(airlines, by = "carrier")
View(joined_flights)

flights %>% 
  select(carrier, flight)

flights_no_year <- flights %>% select(-year)

flights_reorder <- flights %>% 
  select(year, month, day, hour, minute, time_hour, everything())
glimpse(flights_reorder)

flights %>% select(starts_with("a"))
flights %>% select(ends_with("delay"))
flights %>% select(contains("time"))

flights_time_new <- flights %>% 
  select(dep_time, arr_time) %>% 
  rename(departure_time = dep_time, arrival_time = arr_time)
glimpse(flights_time_new)

named_dests %>% top_n(n = 10, wt = num_flights)

named_dests  %>% 
  top_n(n = 10, wt = num_flights) %>% 
  arrange(desc(num_flights))
```


### SUMMARY "
"filter()	Pick out a subset of rows
summarize()	Summarize many values to one using a summary statistic function like
mean(), median(), etc.
group_by()	Add grouping structure to rows in data frame. Note this does not
change values in data frame, rather only the meta-data
mutate()	Create new variables by mutating existing ones
arrange()	Arrange rows of a data variable in ascending (default) or descending order
inner_join()	Join/merge two data frames, matching rows by a key variable"

### CHPT 4

```{r}
dem_score <- read_csv("https://moderndive.com/data/dem_score.csv")
dem_score

?pivot_wider

# to split variable in data
#separate(rate, into = c("cases", "population"))

# to unite data
#unite(new, century, year)
```






