---
title: "1,5 AG Data Analysis Wk 6"
output: html_document
date: '2022-07-27'
---

## Packages
```{r message=FALSE}
library(dplyr)
library(gridExtra)
library(readr)
library(readxl)
library(tidyverse)
library(here)
library(skimr)
library(janitor)
library(ggplot2)
```

## Load Data

```{r message=FALSE, warning=FALSE}
accord_key <- read_csv(paste("accord_key.csv", sep = ""))
activitystatus <- read_csv(paste("activitystatus.csv", sep = ""),
                           col_types = "dcdcdcddd")

bloodpressure <- read_csv(paste("bloodpressure.csv", sep = ""))
hba1c <- read_csv(paste("hba1c.csv", sep = ""))
otherlabs <- read_csv(paste("otherlabs.csv", sep = ""))

cvdoutcomes <- read_csv("cvdoutcomes.csv")
hypoglycemia_events <- read_csv(paste("hypoglycemiaevents.csv", sep = ""))
hypoglycemia_time1st <- read_csv(paste("hypoglycemiatime1st.csv", sep = ""))
sae <- read_csv(paste("sae.csv", sep = ""))

```


1,5 AG data and linkfile. Reading in the data and editing variables. 
```{r message=FALSE, warning=FALSE}


AG <- read_xlsx(paste("UCLA Samples-2022-07-19 Summary.xlsx"), 
                sheet = "Summary", skip = 7) %>% 
  rename(BSI_ID = `Sample ID`)
linkfile <- read_csv("zhou_7816_linkfile.csv")

AG <- AG %>% 
  full_join(linkfile, by = "BSI_ID") %>% 
  filter(!is.na(MASKID) ) %>% 
  rename(MaskID = MASKID,
         Box = `R2020 000036`)

BatchIDs <- c("R2020 000036", "R2020 000037", "R2020 000038", "R2020 000039", 
              "R2020 000040", "R2020 000041", "R2020 000042", "R2020 000043")
index = c(1, which(AG$Box %in% BatchIDs), 4001)
BatchID_col = rep(BatchIDs, diff(index))
AG <- AG %>% mutate(BatchID = BatchID_col) %>%
  relocate(BatchID, .before = Box) %>%
  relocate(MaskID, .after = BSI_ID)

AG_summary <- AG %>% group_by(Box)
```


Adding the hba1c after 12 months data with 1,5 AG
```{r warning=FALSE}
hba1c_12month <- hba1c %>%
  filter(Visit == "F12")
AG_hba1c <- AG %>% 
  left_join(hba1c_12month, by = "MaskID") 
```
```{r echo=FALSE, warning=FALSE}
AG %>% 
  left_join(hba1c_12month, by = "MaskID") 
```


## Plotting

```{r warning=FALSE}
AG %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter()
```
```{r echo=FALSE, warning=FALSE}
AG %>%
  filter(Mean < 50) %>%
  ggplot(aes(x = BatchID, y = Mean)) +
  geom_jitter()
```

Scatter plot for every BatchID
```{r warning=FALSE}
AG %>%
  filter(Mean < 1000, Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  group_by(BatchID) %>%
  ggplot(aes(x = Box, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```
```{r echo=FALSE, warning=FALSE}
AG %>%
  filter(Mean < 50, Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  group_by(BatchID) %>%
  ggplot(aes(x = Box, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ BatchID)
```



### Box Plots

Box Plot for Box 1-7 of all batches
```{r warning=FALSE}
AG %>%
  filter(Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = Box, y = Mean)) +
  geom_boxplot()
```

Box Plot of every box per BatchID
```{r warning=FALSE}
AG %>%
  group_by(BatchID) %>%
 filter(Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = Box, y = Mean)) +
  geom_boxplot() +
  facet_wrap(~ BatchID)
```
```{r warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = Box, y = hba1c)) +
  geom_boxplot() +
  facet_wrap(~ BatchID)

```

hba1c vs. 1,5 AG graph
```{r warning=FALSE}
AG_hba1c %>%
  filter(Mean < 40) %>%
  ggplot(aes( x = hba1c, y = Mean)) +
  geom_jitter() +
  labs(title = "", 
     subtitle = "",
     caption = "",
     x = "Glycated Hemoglobin (%)",
     y = "1,5 Anhyrdoglucitol (µg/m") +
  geom_abline(intercept = 40, slope = -3.75)

```

For each individual batch
```{r warning=FALSE}
AG_hba1c %>%
  filter(Mean < 1000, Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  group_by(BatchID) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
    facet_wrap(~ BatchID)
```

```{r echo=FALSE, warning=FALSE}

AG_hba1c %>%
  filter(Mean < 50, Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  group_by(BatchID) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  geom_abline(intercept = 40, slope = -3.75) +
  facet_wrap(~ BatchID)

```



Scatter plot for each individual Box within each BatchID

### Batch 36
```{r warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Mean < 70,  BatchID %in% c("R2020 000036")) %>%
  group_by(Box) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ Box)
```

### Batch 37

```{r warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Mean < 70,  BatchID %in% c("R2020 000037"), Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ Box)
```

### Batch 38

```{r message=TRUE, warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Mean < 70,  BatchID %in% c("R2020 000038"), Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ Box)
```

### Batch 39 

```{r warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Mean < 70,  BatchID %in% c("R2020 000039"), Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ Box)
```

### Batch 40
```{r warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Mean < 70,  BatchID %in% c("R2020 000040"), Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ Box)
```

### Batch 41
```{r warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Mean < 70,  BatchID %in% c("R2020 000041"), Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ Box)
```

### Batch 42

```{r warning=FALSE}
AG_hba1c %>%
  group_by(BatchID) %>%
  filter(Mean < 70,  BatchID %in% c("R2020 000043"), Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  ggplot(aes(x = hba1c, y = Mean)) +
  geom_jitter() +
  facet_wrap(~ Box)
```




## Summary Statistics]

Total data 
```{r}
AG_hba1c %>%
  filter(Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  group_by(Box) %>%
  summarize(min = min(Mean, na.rm = TRUE),
            q1 = quantile(Mean, 0.25, na.rm = TRUE),
            median = quantile(Mean, 0.5, na.rm = TRUE),
            q3 = quantile(Mean, 0.75, na.rm = TRUE),
            max = max(Mean, na.rm = TRUE),
            mean = mean(Mean, na.rm = TRUE),
            sd = sd(Mean, na.rm = TRUE),)
```

Per Box within every BatchID
```{r}
AG_hba1c %>%
  filter(Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  group_by(Box, BatchID) %>%
  summarize(min = min(Mean, na.rm = TRUE),
            q1 = quantile(Mean, 0.25, na.rm = TRUE),
            median = quantile(Mean, 0.5, na.rm = TRUE),
            q3 = quantile(Mean, 0.75, na.rm = TRUE),
            max = max(Mean, na.rm = TRUE),
            mean = mean(Mean, na.rm = TRUE),
            sd = sd(Mean, na.rm = TRUE),)
```

Per BatchID
```{r}
AG %>%
#  filter(Box %in% c("Box 1", "Box 2", "Box 3", "Box 4", "Box 5", "Box 6", "Box 7")) %>%
  group_by(BatchID) %>%
  summarize(min = min(Mean, na.rm = TRUE),
            q1 = quantile(Mean, 0.25, na.rm = TRUE),
            median = quantile(Mean, 0.5, na.rm = TRUE),
            q3 = quantile(Mean, 0.75, na.rm = TRUE),
            max = max(Mean, na.rm = TRUE),
            mean = mean(Mean, na.rm = TRUE),
            sd = sd(Mean, na.rm = TRUE),)

```

For all the data
```{r}
AG %>%
  summarize(min = min(Mean, na.rm = TRUE),
            q1 = quantile(Mean, 0.25, na.rm = TRUE),
            median = quantile(Mean, 0.5, na.rm = TRUE),
            q3 = quantile(Mean, 0.75, na.rm = TRUE),
            max = max(Mean, na.rm = TRUE),
            mean = mean(Mean, na.rm = TRUE),
            sd = sd(Mean, na.rm = TRUE),)
```


