---
title: "Week 9"
output: html_document
date: '2022-08-25'
---


###Laoading Packages
```{r}
library(tidyverse)
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(readr)       # for importing data
#library(broom.mixed) # for converting bayesian models to tidy tibbles
library(dotwhisker)  # for visualizing regression results
library(skimr)
library(nycflights13)
```

### 1. Build a model 
#L##oading in data 
```{r}
urchins <-
  read_csv("https://tidymodels.org/start/models/urchins.csv") %>% 
  setNames(c("food_regime", "initial_volume", "width")) %>% 
  mutate(food_regime = factor(food_regime, levels = c("Initial", "Low", "High")))

```


### Plotting the data 
```{r}

ggplot(urchins,
       aes(x = initial_volume, 
           y = width, 
           group = food_regime, 
           col = food_regime)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = .7)


```

```{r}
urchins %>% 
  group_by(food_regime) %>% 
  summarize(med_vol = median(initial_volume))

```


```{r}
ggplot(urchins,
       aes(initial_volume, width)) +      
  geom_jitter() +                         
  geom_smooth(method = lm, se = FALSE) +                   
  labs(x = "Volume", y = "Width")         

```


###2. Process your data with recipes

####loading the data
```{r}
set.seed(123)

flight_data <- 
  flights %>% 
  mutate(
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    arr_delay = factor(arr_delay),
    date = lubridate::as_date(time_hour)
  ) %>% 
  inner_join(weather, by = c("origin", "time_hour")) %>% 
  select(dep_time, flight, origin, dest, air_time, distance, 
         carrier, date, arr_delay, time_hour) %>% 
  na.omit() %>% 
  mutate_if(is.character, as.factor)

```


###We can see that about 16% of the flights in this data set arrived more than 30 minutes late.
```{r}

flight_data %>% 
  count(arr_delay) %>% 
  mutate(prop = n/sum(n))

glimpse(flight_data)

flight_data %>% 
  skimr::skim(dest, carrier) 

```


### Fix the random numbers by setting the seed 
### This enables the analysis to be reproducible when random numbers are used 

```{r}
set.seed(222)


flight_data %>% 
  distinct(date) %>% 
  mutate(numeric_date = as.numeric(date)) 

```


##Estimates area under the curve
```{r}
#flights_aug %>% 
#  roc_auc(truth = arr_delay, .pred_late)
```

```{r}
#> # A tibble: 1 × 3
#>   .metric .estimator .estimate
#>   <chr>   <chr>          <dbl>
#> 1 roc_auc binary         0.764

```


### 3 --------
```{r}
data(cells, package = "modeldata")
cells

cells %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))

set.seed(123)
cell_split <- initial_split(cells %>% select(-case), 
                            strata = class)
cell_train <- training(cell_split)
cell_test  <- testing(cell_split)

```

```{r}
nrow(cell_train)
#> [1] 1514
nrow(cell_train)/nrow(cells)
#> [1] 0.7498762

# training set proportions by class
cell_train %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))
#> # A tibble: 2 × 3
#>   class     n  prop
#>   <fct> <int> <dbl>
#> 1 PS      975 0.644
#> 2 WS      539 0.356
```


###test set proportions by class
```{r}
cell_test %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))
#> # A tibble: 2 × 3
#>   class     n  prop
#>   <fct> <int> <dbl>
#> 1 PS      325 0.644
#> 2 WS      180 0.356
```

```{r}
rf_mod <- 
  rand_forest(trees = 1000) %>% 
  set_engine("ranger") %>% 
  set_mode("classification")
```


```{r}
data(cells, package = "modeldata")
cells

set.seed(123)
cell_split <- initial_split(cells %>% select(-case), 
                            strata = class)
cell_train <- training(cell_split)
cell_test  <- testing(cell_split)

```

```{r}

tune_spec <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()
  ) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")

tune_spec
```

```{r}

tree_grid <- grid_regular(cost_complexity(),
                          tree_depth(),
                          levels = 5)
```

```{r}
tree_grid %>% 
  count(tree_depth)

set.seed(234)
cell_folds <- vfold_cv(cell_train)

```

```{r}

set.seed(345)

tree_wf <- workflow() %>%
  add_model(tune_spec) %>%
  add_formula(class ~ .)
```

```{r}
tree_res <- 
  tree_wf %>% 
  tune_grid(
    resamples = cell_folds,
    grid = tree_grid
  )

tree_res %>% 
  collect_metrics()

```

```{r}
tree_res %>%
  collect_metrics() %>%
  mutate(tree_depth = factor(tree_depth)) %>%
  ggplot(aes(cost_complexity, mean, color = tree_depth)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)



```



